A Hedonic Price Analysis of Traffic Noise in the Twin Cities Housing Market

### 2013-08-30 reports/Draft.Rnw
worked on the Dakota County table and future research

An important first step towards such an analysis is to examine real estate price fluctuations as a function of noise levels using hedonic regression. \citet{Nelson1982} reviewed early work in this area, but many early studies were limited by small sample sizes and narrow spatial scopes. A review by \citet{Bateman2001} found that the negative eﬀect of traﬃc noise ranged from 0.08 to 2.22 percent per dB with a mean around 0.55 percent per dB, while \citet{Navrud2002} found an average decrease of 0.64 percent per dB. More recent work has tended to focus on the European experience with noise, especially the United Kingdom \citep{Day2007, Blanco2011}, Sweden \citep{Wilhelmsson2000, Andersson2010}, Netherlands \citep{Theebe2004a}, Spain \citep{MarmolejoDuarteCarlos;GonzalezTamez2009}, and Switzerland \citep{Baranzini2010}.

Work examining the negative effects of noise in the United States have tended to concentrate on noise from airports \citep{Espey2000, McMillen2004, Cohen2008a} or indirect effects of traffic noise, such as proximity to highways \citep{Matthews2007, Chernobai2009, Li2012}, or traffic counts \citep{Highes1992, Larsen2012}. the relationship between road traffic noise levels and real estate prices has rarely been explored in the United States, even though the former is perhaps the greatest source of noise in residential neighborhoods (XXX).  One major reason for the thinness of the scholarly literature on this topic in the US lies in the difficulty of modeling noise propagation over the landscape.

@@

In this paper, we seek to contribute to the hedonic literature by examining the relationship between real estate prices and variation in road traffic noise exposure.  We approach this issue by means of an empirical case study of the St.\ Paul, Minnesota urban area.  In particular, we construct a locally flexible model that allows for a non-stationary relationship between house prices and our explanatory variables over geography and time. Such Locally Weighted Regression (LWR) models have become increasingly common in published work and have been shown to be more accurate than spatial autoregressive models with misspecified weights matrices.

Our work makes three contributions to the field. First, we utilize new estimates of traffic noise that has never before been used in the hedonic literature. The data explicitly model noise propagation as a function of nearby buildings and landcover, as well as present the results at a fine spatial resolution (10m x 10m). Such estimates, combined with sptially explicit modeling techniques allow us to make our second contribution to the literature. We present one of the first estimates of how the impact of traffic noiseeee varies over space. We find that while the impact of other important control variables in the hedonic function vary over space, the impact of traffic noise does not. Lastly, we present the first estimates of how the Great Recession has affected the hedonic price of traffic noise.

@@

Taken in aggregate, the results in Table \ref{tab:globalRegression} suggest that the hedonic price function may vary over space and time. While methods exist for parameterizing this variation (such as spatial expansion as suggested by \citet{Casetti1974}), we have no a priori knowledge of how to parameterize the variation over space and time. As such, we turn to a semi-parametric form of hedonic regression, a flexible modeling approach which lets the data reveal how relationships vary, rather than specifying them beforehand.

@@

\begin{table}[bp]
\caption{Hedonic Regression Results (dependent variable = $ln(\textrm{Sale Price})$)} \label{tab:mainresults} 
\makebox[\linewidth]{
\small
\begin{tabular}{lrrrrrr} \\
 & \multicolumn{2}{c}{Global Model} & \multicolumn{4}{c}{Locally Weighted Regression Model} \\
 & \multicolumn{1}{c}{$\hat{\beta}$} & \multicolumn{1}{c}{$\hat{\sigma}_{\hat{\beta}}$} & mean $\hat{\beta}$ & \multicolumn{1}{c}{$\sigma_{\hat{\beta}}$} & (5) & (6) \\ 
  \hline
(Intercept) & 8.4E+00 & 1.3E-01 & 2.9E+00 & 1.5E+01 & 0.64 & 0.00 \\ 
  Traffic Noise & -2.6E-03 & 1.6E-04 & -2.5E-03 & 3.0E-03 & 0.47 & 0.67 \\ 
  House Size & 3.0E-04 & 2.7E-06 & 2.6E-04 & 9.1E-05 & 1.00 & 0.00 \\ 
  Lot Size & 2.6E-01 & 1.5E-02 & 4.0E-01 & 3.7E-01 & 0.67 & 0.00 \\ 
  Year of Construction & 1.2E-03 & 6.5E-05 & 4.5E-03 & 5.4E-03 & 0.79 & 0.00 \\ 
  Owner Occupancy & 2.7E-02 & 3.2E-03 & 2.3E-02 & 6.4E-02 & 0.37 & 1.00 \\ 
  Elementary Test Scores & 1.5E-03 & 1.2E-04 & 2.4E-04 & 3.1E-02 & 0.30 & 0.00 \\ 
  Median Income & 3.3E-06 & 7.7E-08 & 9.6E-07 & 8.1E-06 & 0.35 & 0.00 \\ 
  distance to CBD & 1.3E-05 & 6.3E-07 & 1.1E-05 & 1.0E-04 & 0.30 & 0.00 \\ 
  `` nearest Park & -3.7E-07 & 1.1E-06 & -1.6E-05 & 1.3E-04 & 0.32 & 0.00 \\ 
  `` nearest Lake & 2.5E-05 & 2.1E-06 & -2.8E-05 & 1.4E-04 & 0.35 & 0.00 \\ 
  `` nearest Shopping Center & -5.3E-06 & 1.3E-06 & 1.5E-05 & 8.8E-05 & 0.29 & 0.00 \\  
  location of sale & \multicolumn{2}{c}{city fixed effects} & \multicolumn{4}{c}{\footnotesize city fixed effects and nearest 200 sales} \\
  timing of sale & \multicolumn{2}{c}{year fixed effects} & \multicolumn{4}{c}{within last 12 months} \\
 \hline
  GCV score & \multicolumn{2}{c}{0.041} & \multicolumn{2}{c}{0.027}  &  & \\
  Moran's I statistic & \multicolumn{2}{c}{0.199} & \multicolumn{2}{c}{0.012} & & \\
\end{tabular}
}
 \caption*{\small (5) displays the proportion of $\hat{\beta}_{LWR}$ that can reject the null hypthesis of $\beta =0$.\\ (6) displays the proportion of Monte Carlo simulations with larger standard deviations of the LWR coefficients than $\sigma_{\hat{\beta}}$}
\end{table}

@@

\subsection{Does Space Matter?}
We believe the use of a LWR estimation procedure is appropriate after conducting two different Monte Carlo simulations with the data. In each simulation we resampled (without replacement) the location of each house in the study and then re-estimated the LWR model with the resampled data. In the first simulation we estimated the LWR model using bandwidths ranging from the nearest 50 to 10,000 sales as before. In 100 consecutive simulations, the smallest Generalized Cross-Validation score was obtained at the largest bandwidth. The smallest of these GCV scores was 0.041, just as was the case with the global (non-LWR) model. In other words, the LWR model never performed better than the global model when the locations of the data were randomized. This seems to be strong evidence that the increased ability to predict house prices with a local model is not due to chance.

@@

\subsection{Do the Regression Coefficients Vary over Time}

We know that house prices fell dramatically during and after the Great Recession of 2008-09. We know less about the patterns of change in the hedonic price functions during and after the fall in prices. For instance, did all house prices simply drop (mostly affecting the intercept term of the hedonic function), or did the implicit price of structural characteristics like the size of the house or its lot decrease, or perhaps the value of environmental amenities disproportionately dropped? This section presents some of the first evidence of how the hedonic price function for housing changed over time surrounding the Great Recession of 2008-09. We ask whether the LWR coefficients exhibit a temporal pattern.

Specifically, we create two new variables, ``\texttt{months since September 2008}'' and a dummy variable ``\texttt{post}'' for all sales after September 2008. We then regress our LWR coefficients on these two variables and their interaction term as shown in equation \eqref{eq:LWRinteraction}. 
\begin{equation}\label{eq:LWRinteraction}
\hat{\beta}_{LWR} = \alpha _0 + \alpha _1 \texttt{*month} + \alpha _2  \texttt{*post} + \alpha _3 \texttt{*post*month} + \epsilon
\end{equation}

The results of these regressions will help us understand the patterns of change in the hedonic price of different characteristics. In particular, significant values for $\hat{\alpha}_2$ suggest that there was a structural break in the LWR coefficients post September 2008, and significant values for $\hat{\alpha}_3$ will imply a difference in the temporal trend of the LWR coefficients.

Note that the tables have not yet been formatted.

\subsubsection{Traffic Noise}

We know of no published research that shows how the hedonic price estimates of traffic noise vary over time surrounding the Great Recession. In this section we show that accounting for the month of the house sale significantly predicts to the estimated LWR coefficients for traffic noise. The linear regression results presented in Table \ref{tab:betaMAX} show that there is a significant negative shock to the estimated LWR coefficients for traffic noise after September 2008. Additionally, the coefficients also begin to a significant negative trend.
\begin{table}[h]
\caption{Regression Results: Dependent Variable = Traffic Noise LWR Coefficients}\label{tab:betaMAX}
\begin{verbatim}
(Intercept) -1.781e-03  4.100e-05 -43.436  < 2e-16 ***
month        6.982e-06  2.005e-06   3.482 0.000498 ***
Post        -7.998e-04  7.209e-05 -11.094  < 2e-16 ***
month:Post  -8.868e-05  4.484e-06 -19.776  < 2e-16 ***
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

@@

The results in Table \ref{tab:betaMAX} contradict the a priori assumption that environmental amenities will ``matter'' less during and after the Great Recession. In fact, these results suggest that the penalty for homes exposed to higher levels of traffic noise increased. That is, the negative hedonic coefficients on traffic noise got more negative after September 2008. It should be noted that our model can explain almost 10 percent of the variation in traffic noise LWR coefficient estimates.

\subsubsection{House Size}

In this section we show how the hedonic coefficients on house size (as measured in square feet) change over time. Table \ref{tab:betaSQFT} shows that the LWR coefficients on house size were positive and trending up before September 2008, then the coefficients drop significantly in both absolute value and their time trend. These results are consistent with the belief that the Great Recession would reduce consumer's marginal willingness to pay for increases in the size of a house. It should be noted that the model presented in Table \ref{tab:betaSQFT} explains less than one percent of the variation in our house size LWR coefficients.

\begin{table}[h]
\caption{Regression Results: Dependent Variable = House Size LWR Coefficients}\label{tab:betaSQFT}
\begin{verbatim}
              Estimate Std. Error t value Pr(>|t|)    
(Intercept)  2.741e-04  1.317e-06 208.097  < 2e-16 ***
month        8.742e-07  6.441e-08  13.573  < 2e-16 ***
Post        -1.678e-05  2.316e-06  -7.245 4.43e-13 ***
month:Post  -5.989e-07  1.440e-07  -4.157 3.23e-05 ***
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

Residual standard error: 9.04e-05 on 31744 degrees of freedom
Multiple R-squared: 0.006092,  Adjusted R-squared: 0.005998 
F-statistic: 64.85 on 3 and 31744 DF,  p-value: < 2.2e-16 
\end{verbatim}
\end{table}

\subsubsection{Lot Size}

Table \ref{tab:betaACRES} shows interesting results that are qualitatively different from our previous results. The post-September 2008 dummy variable coefficient is only marginally significant and small in absolute value (when compared to the regression Intercept term). There is no discernable change in the temporal trend of the lot size LWR coefficients. The model presented in Table \ref{tab:betaACRES} explains less than a half of one percent of the variation in lot size LWR coefficients.

\begin{table}[h]
\caption{Regression Results: Dependent Variable = Lot Size LWR Coefficients}\label{tab:betaACRES}
\begin{verbatim}
(Intercept)  0.3828690  0.0054089  70.785  < 2e-16 ***
month       -0.0015113  0.0002645  -5.713 1.12e-08 ***
Post         0.0190349  0.0095117   2.001   0.0454 *  
month:Post  -0.0001358  0.0005916  -0.229   0.8185    
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

Residual standard error: 0.3713 on 31744 degrees of freedom
Multiple R-squared: 0.002806,  Adjusted R-squared: 0.002711 
F-statistic: 29.77 on 3 and 31744 DF,  p-value: < 2.2e-16 
\end{verbatim}
\end{table}

\subsubsection{Intercept}
We were also interested in how the LWR regression intercept changed over time. For instance, perhaps the house price drops of the Great Recession were manifested primarily through reduced intercept terms. Table \ref{tab:betaINT} is not consistent with that hypothesis. In fact, our results show that the estimated LWR intercept terms are on average significantly higher post-September 2008.
\begin{table}[h]
\caption{Regression Results: Dependent Variable = LWR Intercept Estimates}\label{tab:betaINT}
\begin{verbatim}
(Intercept)  1.575488   0.213779   7.370 1.75e-13 ***
month       -0.043836   0.010455  -4.193 2.76e-05 ***
Post         2.845774   0.375939   7.570 3.84e-14 ***
month:Post   0.007397   0.023383   0.316    0.752    
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

Residual standard error: 14.67 on 31744 degrees of freedom
Multiple R-squared: 0.003282,  Adjusted R-squared: 0.003188 
F-statistic: 34.84 on 3 and 31744 DF,  p-value: < 2.2e-16  
\end{verbatim}
\end{table}

@@

Some researchers, such as \citet{Theebe2004a} conclude that the impact of traffic noise is non-linear in the level of the noise, while others, such as \citet{Huang;Palmquist2001} report a constant impact per decibel increase. In this section we compare the mean LWR noise coefficient estimates across noise levels to look for non-linearities in the impact of traffic noise.

@@

Table \ref{tab:LWRnonlinear} estimates the predicted LWR noise coefficient as a linear function of the noise level in decibels and whether or not the sale took place before or after September 2008. The table reveals no statistically measurable linear relationship between the noise coefficient for sales before September 2008 and the level of the noise (that is, the marginal effect is no different at low levels of noise than at high levels of noise). The positive and significant interaction term suggests that the impact may be different for high vs. low levels of noise post-September 2008, but the difference in the predicted marginal effects is so small, that it works about to be a roughly \$25 difference for a \$300,000 house at 50 vs. 75 dB.

@@

Our work suffers from at least two noticeable weaknesses. First, we are missing some important structural variables that are commonly used in the real estate hedonic literature. To the extent that structural variables like the number of bedrooms, bathrooms, garage size, or construction quality covary with other variables in our dataset, our regression coefficients will suffer from omitted variable bias. We are somewhat comforted, however, because we do have some additional variables for a subset of our data. We have the number of bedrooms, bathrooms, and garage area for our houses located within Dakota County. Our analysis (Models 1, 2, and 3 described earlier) was repeated on this geographic subset of the data and the results were compared to the estimates obtained from the same analysis when these additional structural variables were omitted. We found strikingly similar estimates of the impact of traffic noise.

\begin{table}[ht]
\begin{center}
\caption{Mean and (Standard Deviation) LWR Noise Coefficients in Dakota County both with and without the additional structural variables available only for the county.}\label{tab:Dak}
\begin{tabular}{lll}
 & \multicolumn{2}{c}{Additional variables?} \\
Locally Weighted Regression Model & \multicolumn{1}{c}{no} & \multicolumn{1}{c}{yes} \\ \hline
Model (1) = Structural Variables & -0.00149 & -0.00147 \\ 
   & (0.00226) & (0.00309) \\ 
Model (2) = (1) + Locational Variables & -0.00136 & -0.00140 \\ 
   & (0.00184) & (0.00218) \\ 
Model (3) = (2) + City Fixed Effects & -0.00140 & -0.00137 \\ 
   & (0.00219) & (0.00176) \\ 
\end{tabular}
\end{center}
\end{table}

Table \ref{tab:Dak} shows that the traffic noise coefficients estimated by our LWR models are similar regardless of whether the additional structural variables were included or not. Welch two-sample t-tests fail to reject the null hypothesis of zero means across our three models (comparing LWR estimates with vs. without the additional structural variables included in the model). Paired t-tests find differences for Model 2 and 3, but the estimated differences are zero to four decimal places and in one case the noise impact estimates with the additional variables are slightly larger and in the other case they are slightly smaller. We also conducted simple linear egressions of the noise coefficients without the additional structural variables on the noise coefficients obtained with the additional structural variables. In all three cases the intercept estimates were close zero with slopes almost exactly equal to one and each also had $R^2$ values over 0.7.

The timing of our independent variable collection is also a potential problem. While our house sales data are collected over the course of six years, some of our other variables were collected at specific points in time and assumed to be constant over the study period. In particular, the traffic noise estimates taken from \citet{Nega2012} are for the year 2007. To the extent that traffic flows and composition signicantly changed over time, our traffic noise variables may be inaccurate for those time periods. For instance, the US Department of Transportation reports that total vehicle miles travelled decreased by up to 2 percent year-on-year during the Great Recession. Thus, our noise estimates may be inaccurate for later time periods and this may bias our estimates of the impact of noise on house values. Future work may better estimate the impact of traffic noise over time by generating noise data to reflect changing traffic patterns.

What other weaknesses should we discuss?

We didn't explicitly model spatial autocorrelation in the error terms.

We might want to further reduce the number of parameters to estimate through the use of either mixed-LWR or running separate regressions for submarkets determined through hierarchical clustering like \citet{Day2003}.

>>>>

An important first step towards such an analysis is to examine real estate price fluctuations as a function of noise levels using hedonic regression. \citet{Nelson1982} reviewed early work in this area, but many early studies were limited by small sample sizes and narrow spatial scopes. A review by \citet{Bateman2001} found that the negative effect of traffic noise ranged from 0.08 to 2.22 percent per dB with a mean around 0.55 percent per dB, while \citet{Navrud2002} found an average decrease of 0.64 percent per dB. More recent work has tended to focus on the European experience with noise, especially the United Kingdom \citep{Day2007, Blanco2011}, Sweden \citep{Wilhelmsson2000, Andersson2010}, Netherlands \citep{Theebe2004a}, Spain \citep{MarmolejoDuarteCarlos;GonzalezTamez2009}, and Switzerland \citep{Baranzini2010}.

The relationship between road traffic noise levels and real estate prices has rarely been explored in the United States, even though the former is perhaps the greatest source of noise in residential neighborhoods (XXX). Work examining the negative effects of noise in the US has instead tended to concentrate on noise from airports \citep{Espey2000, McMillen2004, Cohen2008a} or indirect effects of traffic noise, such as proximity to highways \citep{Matthews2007, Chernobai2009, Li2012}, or traffic counts \citep{HughesJr.1992, Larsen2012}.  One major reason for the thinness of the scholarly literature on this topic in the US lies in the difficulty of modeling noise propagation over the landscape.

@@

In this paper, we seek to contribute to the hedonic literature by examining the relationship between real estate prices and variation in road traffic noise exposure.  We approach this issue by means of an empirical case study of the St.\ Paul, Minnesota urban area.  In particular, we construct a locally flexible model that allows for a non-stationary relationship between house prices and our explanatory variables over geography and time. Such Locally Weighted Regression (LWR) models have become increasingly common in published work \citep[see][]{MarmolejoDuarteCarlos;GonzalezTamez2009, Carruthers2010, Sunding2010, Nappi-Choulet2011} and have been shown to be more accurate than spatial autoregressive models with misspecified weights matrices \citep{McMillen2012}.

Our work makes three contributions to the field. First, we utilize new estimates of traffic noise that has never before been used in the hedonic literature. The data explicitly model noise propagation as a function of nearby buildings and landcover, as well as present the results at a fine spatial resolution (10m x 10m). Such estimates, combined with sptially explicit modeling techniques allow us to make our second contribution to the literature. We present one of the first non-parametric estimates of how the impact of traffic noise varies over space within an urban area in the United States. Contrary to previous work, we find that while the impact of other important control variables in the hedonic function vary over space, the impact of traffic noise does not. Lastly, we present the first estimates of how the Great Recession has affected the hedonic price of traffic noise.

@@

Taken in aggregate, the results in Table \ref{tab:globalRegression} suggest that the hedonic price function may vary over space and time. While methods exist for parameterizing this variation \citep[such as spatial expansion as suggested by][]{Casetti1972}, we have no a priori knowledge of how to parameterize the variation over space and time. As such, we turn to a semi-parametric form of hedonic regression, a flexible modeling approach which lets the data reveal how relationships vary, rather than specifying them beforehand.

@@

\begin{table}[bp]
\caption{Hedonic Regression Results (dependent variable = $ln(\textrm{Sale Price})$)} \label{tab:mainresults} 
\makebox[\linewidth]{
\small
\begin{tabular}{lrrrrrr} \\
 & \multicolumn{2}{c}{Global Model} & \multicolumn{4}{c}{Locally Weighted Regression Model} \\
 & \multicolumn{1}{c}{$\hat{\beta}$} & \multicolumn{1}{c}{$\hat{\sigma}_{\hat{\beta}}$} & mean $\hat{\beta}$ & \multicolumn{1}{c}{$\sigma_{\hat{\beta}}$} & (5) & (6) \\ 
  \hline
(Intercept) & 8.4E+00 & 1.3E-01 & 2.9E+00 & 1.5E+01 & 0.64 & 0.00 \\ 
  Traffic Noise & -2.6E-03 & 1.6E-04 & -2.5E-03 & 3.0E-03 & 0.47 & 0.67 \\ 
  House Size & 3.0E-04 & 2.7E-06 & 2.6E-04 & 9.1E-05 & 1.00 & 0.00 \\ 
  Lot Size & 2.6E-01 & 1.5E-02 & 4.0E-01 & 3.7E-01 & 0.67 & 0.00 \\ 
  Year of Construction & 1.2E-03 & 6.5E-05 & 4.5E-03 & 5.4E-03 & 0.79 & 0.00 \\ 
  Owner Occupancy & 2.7E-02 & 3.2E-03 & 2.3E-02 & 6.4E-02 & 0.37 & 1.00 \\ 
  Elementary Test Scores & 1.5E-03 & 1.2E-04 & 2.4E-04 & 3.1E-02 & 0.30 & 0.00 \\ 
  Median Income & 3.3E-06 & 7.7E-08 & 9.6E-07 & 8.1E-06 & 0.35 & 0.00 \\ 
  distance to CBD & 1.3E-05 & 6.3E-07 & 1.1E-05 & 1.0E-04 & 0.30 & 0.00 \\ 
  `` nearest Park & -3.7E-07 & 1.1E-06 & -1.6E-05 & 1.3E-04 & 0.32 & 0.00 \\ 
  `` nearest Lake & 2.5E-05 & 2.1E-06 & -2.8E-05 & 1.4E-04 & 0.35 & 0.00 \\ 
  `` nearest Shopping Center & -5.3E-06 & 1.3E-06 & 1.5E-05 & 8.8E-05 & 0.29 & 0.00 \\  
  location of sale & \multicolumn{2}{c}{city fixed effects} & \multicolumn{4}{c}{\footnotesize city fixed effects and nearest 200 sales} \\
  timing of sale & \multicolumn{2}{c}{year fixed effects} & \multicolumn{4}{c}{within last 12 months} \\
 \hline
  GCV score & \multicolumn{2}{c}{0.041} & \multicolumn{2}{c}{0.027}  &  & \\
  Moran's I statistic & \multicolumn{2}{c}{0.199} & \multicolumn{2}{c}{0.012} & & \\
\end{tabular}
}
 \caption*{\small Column (5) displays the proportion of $\hat{\beta}_{LWR}$ that can reject the null hypthesis of $\beta =0$.\\ Column (6) displays the proportion of Monte Carlo simulations with larger standard deviations of the LWR coefficients than $\sigma_{\hat{\beta}}$}
\end{table}

@@

\subsection{How Does the Hedonic Function Vary over Space?}
We believe the use of a LWR estimation procedure is appropriate after conducting two different Monte Carlo simulations with the data. In each simulation we resampled (without replacement) the location of each house in the study and then re-estimated the LWR model with the resampled data. In the first simulation we estimated the LWR model using bandwidths ranging from the nearest 50 to 10,000 sales as before. In 100 consecutive simulations, the smallest Generalized Cross-Validation score was obtained at the largest bandwidth. The smallest of these GCV scores was 0.041, just as was the case with the global (non-LWR) model. In other words, the LWR model never performed better than the global model when the locations of the data were randomized. This seems to be strong evidence that the increased ability to predict house prices with a local model is not due to chance.

@@

\subsection{Does Traffic Noise Penalty Change over Time?}

Previous research has suggested that the impact of noise can change with time and economic conditions. For instance, \citet{Wilhelmsson2000} found that the traffic noise penalty to be stronger in the 1990s near Stockholm, Sweden compared to the 1980s. \citet{Cohen2009} also found the (airport) noise penalty to be larger in the early 2000s compared to the late 1990s in Atlanta, Georgia. We know that house prices in the United States fell dramatically during and after the Great Recession of 2008-09. We know less about the patterns of change in the hedonic price functions during and after the fall in prices. \citet{Cho2011b} found evidence from hedonic analysis in the Nashville, Tennessee area that consumers' willingness to pay for environmental amenities (such as proximity to open space and water views) decreased during the recession as compared to the previous economic boom.

In this section we look for evidence of temporal non-stationarity in the hedonic coefficient on traffic noise within our data. Specifically, we create two new variables, ``\texttt{months since September 2008}'' and a dummy variable ``\texttt{post}'' for all sales after September 2008. We then regress our LWR noise coefficients on these two variables and their interaction term as shown in equation \eqref{eq:LWRinteraction}, 
\begin{equation}\label{eq:LWRinteraction}
\hat{\beta}_{LWR} = \alpha _0 + \alpha _1 \texttt{*month} + \alpha _2  \texttt{*post} + \alpha _3 \texttt{*post*month} + \epsilon.
\end{equation}
The results of this regression will help us understand the patterns of change in the implicit price of traffic noise. In particular, $\hat{\alpha}_1 \neq 0$ suggests a linear temporal trend in the marginal willingness to pay to avoid traffic noise, while $\hat{\alpha}_2 \neq 0$ suggests that there was a structural break after September 2008, and $\hat{\alpha}_3 \neq 0$ will imply a change in the temporal trend after September 2008. The linear regression results presented in Table \ref{tab:betaMAX} show that there is a significant negative shock to the estimated LWR coefficients for traffic noise after September 2008. Additionally, the coefficients also begin to a significant negative trend.
\begin{table}[h]
\caption{Regression Results: Dependent Variable = Traffic Noise LWR Coefficients}\label{tab:betaMAX}
\begin{verbatim}
            Estimate Std. Error t value Pr(>|t|)   
(Intercept) -1.7e-03   4.1e-05   -43.4   < 2e-16 ***
Month        6.9e-06   2.0e-06     3.5  0.000498 ***
Post        -8.0e-04   7.2e-05   -11.1   < 2e-16 ***
Month*Post  -8.9e-05   4.5e-06   -19.7   < 2e-16 ***
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

@@

The results in Table \ref{tab:betaMAX} contradict the a priori expectation that environmental amenities will ``matter'' less during and after the Great Recession. In fact, these results suggest that the penalty for homes exposed to higher levels of traffic noise increased. That is, the negative hedonic coefficients on traffic noise got more negative after September 2008. It should be noted that our model can explain almost 10 percent of the variation in traffic noise LWR coefficient estimates.

% \subsubsection{House Size}
% 
% In this section we show how the hedonic coefficients on house size (as measured in square feet) change over time. Table \ref{tab:betaSQFT} shows that the LWR coefficients on house size were positive and trending up before September 2008, then the coefficients drop significantly in both absolute value and their time trend. These results are consistent with the belief that the Great Recession would reduce consumer's marginal willingness to pay for increases in the size of a house. It should be noted that the model presented in Table \ref{tab:betaSQFT} explains less than one percent of the variation in our house size LWR coefficients.
% 
% \begin{table}[h]
% \caption{Regression Results: Dependent Variable = House Size LWR Coefficients}\label{tab:betaSQFT}
% \begin{verbatim}
%               Estimate Std. Error t value Pr(>|t|)    
% (Intercept)  2.741e-04  1.317e-06 208.097  < 2e-16 ***
% month        8.742e-07  6.441e-08  13.573  < 2e-16 ***
% Post        -1.678e-05  2.316e-06  -7.245 4.43e-13 ***
% month:Post  -5.989e-07  1.440e-07  -4.157 3.23e-05 ***
% ---
% Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1 
% 
% Residual standard error: 9.04e-05 on 31744 degrees of freedom
% Multiple R-squared: 0.006092,  Adjusted R-squared: 0.005998 
% F-statistic: 64.85 on 3 and 31744 DF,  p-value: < 2.2e-16 
% \end{verbatim}
% \end{table}
% 
% \subsubsection{Lot Size}
% 
% Table \ref{tab:betaACRES} shows interesting results that are qualitatively different from our previous results. The post-September 2008 dummy variable coefficient is only marginally significant and small in absolute value (when compared to the regression Intercept term). There is no discernable change in the temporal trend of the lot size LWR coefficients. The model presented in Table \ref{tab:betaACRES} explains less than a half of one percent of the variation in lot size LWR coefficients.
% 
% \begin{table}[h]
% \caption{Regression Results: Dependent Variable = Lot Size LWR Coefficients}\label{tab:betaACRES}
% \begin{verbatim}
% (Intercept)  0.3828690  0.0054089  70.785  < 2e-16 ***
% month       -0.0015113  0.0002645  -5.713 1.12e-08 ***
% Post         0.0190349  0.0095117   2.001   0.0454 *  
% month:Post  -0.0001358  0.0005916  -0.229   0.8185    
% ---
% Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1 
% 
% Residual standard error: 0.3713 on 31744 degrees of freedom
% Multiple R-squared: 0.002806,  Adjusted R-squared: 0.002711 
% F-statistic: 29.77 on 3 and 31744 DF,  p-value: < 2.2e-16 
% \end{verbatim}
% \end{table}
% 
% \subsubsection{Intercept}
% We were also interested in how the LWR regression intercept changed over time. For instance, perhaps the house price drops of the Great Recession were manifested primarily through reduced intercept terms. Table \ref{tab:betaINT} is not consistent with that hypothesis. In fact, our results show that the estimated LWR intercept terms are on average significantly higher post-September 2008.
% \begin{table}[h]
% \caption{Regression Results: Dependent Variable = LWR Intercept Estimates}\label{tab:betaINT}
% \begin{verbatim}
% (Intercept)  1.575488   0.213779   7.370 1.75e-13 ***
% month       -0.043836   0.010455  -4.193 2.76e-05 ***
% Post         2.845774   0.375939   7.570 3.84e-14 ***
% month:Post   0.007397   0.023383   0.316    0.752    
% ---
% Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1 
% 
% Residual standard error: 14.67 on 31744 degrees of freedom
% Multiple R-squared: 0.003282,  Adjusted R-squared: 0.003188 
% F-statistic: 34.84 on 3 and 31744 DF,  p-value: < 2.2e-16  
% \end{verbatim}
% \end{table}

@@

Some researchers conclude that the semi-elasticity of noise varies with the level of noise \citep{Andersson2010, MarmolejoDuarteCarlos;GonzalezTamez2009, Theebe2004a, Miedema2001, Wilhelmsson2000}, such as a ``threshold''' effect at 70dB \citep{Wilhelmsson2000, Cohen2009}. Other researchers report no evidence of varying impacts at higher or lower noise intensities \citep{Blanco2011, Baranzini2010, Kim2007, Huang;Palmquist2001}. In this section we compare our mean LWR noise coefficient estimates across noise levels to look for threshold effects of traffic noise.

@@

Table \ref{tab:LWRnonlinear} reports the results of a simple linear regression of the LWR noise coefficients on the noise levels and whether or not the sale took place before or after September 2008 (as well as their interaction). The table reveals no statistically measurable linear relationship between the noise coefficient and the noise level for sales before September 2008. That is, the semi-elasticity is no different in our data at low levels of noise than at high levels of noise. The positive and significant interaction term suggests that the impact may be different for high vs. low levels of noise post-September 2008, but the difference in the predicted marginal effects is so small that it works about to be a roughly \$25 difference for a \$300,000 house at 50 vs. 75 dB. Thus, while other researchers report finding impact thresholds at 70dB, we cannot corroborate such findings with our data.

@@

Our work suffers from at least two noticeable weaknesses. First, we are missing some important structural variables that are commonly used in the housing hedonic literature. To the extent that structural variables like the number of bedrooms, bathrooms, garage size, or construction quality covary with other variables in our dataset, our regression coefficients will suffer from omitted variable bias. We are somewhat comforted, however, because we do have some additional variables for a subset of our data. We obtained the number of bedrooms, bathrooms, and garage area for our houses located within Dakota County from the Dakota County Assessor's Office. Our analysis (Models 1, 2, and 3 described earlier) was repeated on this geographic subset of the data and the results were compared to the estimates obtained from the same analysis when these additional structural variables were omitted. We found strikingly similar estimates of the impact of traffic noise.

\begin{table}[ht]
\begin{center}
\caption{A Comparison of the Noise Regression Coefficient Estimates in Dakota County with and without Additional Structural Variables}\label{tab:Dak}
\begin{tabular}{lllc}
 & \multicolumn{2}{p{1.25in}}{Additional Variables Included in LWR?} & \\
Locally Weighted Regression Model & \multicolumn{1}{c}{no} & \multicolumn{1}{c}{yes} & \\ \cline{1-3}
Model (1) = Structural Variables & -0.0015 & -0.0015 & mean\\ 
   & (0.0023) & (0.0031) & (st.\ dev.) \\ 
Model (2) = Model (1) + Locational Variables & -0.0014 & -0.0014  & \\ 
   & (0.0018) & (0.0022) &  \\ 
Model (3) = Model (2) + City Fixed Effects & -0.0014 & -0.0014 & \\ 
   & (0.0022) & (0.0018) & \\ 
\end{tabular}
\end{center}
\end{table}

Table \ref{tab:Dak} shows that the traffic noise coefficients estimated by our LWR models are similar regardless of whether the additional structural variables were included or not. Welch two-sample t-tests fail to reject the null hypothesis of zero mean differences across our three models (comparing LWR estimates with vs. without the additional structural variables included in the model). Paired t-tests find differences for Model 2 and 3, but the estimated differences are zero to four decimal places and in one case the noise impact estimates with the additional variables are slightly larger and in the other case they are slightly smaller. We also conducted simple linear regressions of the noise coefficients without the additional structural variables on the noise coefficients obtained with the additional structural variables for each of our three models. In all cases the intercept estimates were close zero with slopes almost exactly equal to one and each also had $R^2$ values over 0.7, signifying that the hedonic coefficients for the traffic noise variable are unaffected by omitted variable bias from common structural variables. While we cannot confirm that this lack of ommitted variable bias also occurs elsewhere in our data, the results from Dakota County are promising.
